---
title: "Lake Ontario_Harding"
author: "Riley Harding"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This is a brief report on the microbial abundance in Lake Ontario and the possible explanatory variables for the perceived fluctuation in their abundance temporally.

### Load required packages:

```{r packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(harrypotter)
library(ggpubr)
library(knitr)
```

### Load dataframe

```{r dataframe, message=FALSE, warning=FALSE}
sample_and_taxon <- read_csv("~/OneDrive/Documents/Cornell Grad School/Graduate Classes/Workshops/2025_Carpentries_Coding/ontario-report/data/sample_and_taxon.csv")

buoy_data<-read_csv("~/OneDrive/Documents/Cornell Grad School/Graduate Classes/Workshops/2025_Carpentries_Coding/ontario-report/data/buoy_data.csv")
```

This dataframe was created by merging two dataframes together with R base function inner join.

Now we would like to explore what is impacting microbial abundance in Lake Ontario.

```{r explore}
sample_and_taxon %>% 
ggplot()+
  aes(x = temperature, 
      y = cells_per_ml/1000000,
      color = env_group,
      size = chlorophyll,
      shape = env_group)+
  geom_point()+
  labs(x = "Temperature (°C)", 
       y = "Cell Abundance (millions/mL)", 
       title = "What Influences Microbial Abundance in Lake Ontario",
       size = "Chlorophyll (ug/L)",
       color = "Environmental Group",
       shape = "Environmental Group")
```

From this figure, there seems to be an influence of depth and temperature on the relative abundance of microbial organisms. 

Next, let's look at another dataset that summarizes lake temperatures from four buoys across Lake Ontario.


```{r buoy plot}
buoy_data %>% 
  ggplot()+
  aes(x = day_of_year,
      y = temperature,
      color = depth,
      group = sensor)+
  geom_line()+
  facet_wrap(~buoy, scales = "free_y")+
  labs(x = "Day of the Year (Julian)",
       y = "Temperature (°C)",
       color = "Lake Depth (m)",
       title = "Buoy Temperature Data Across Lake Ontario")
```

There appears to be differences in the distribution of lake temperatures among the four buoys. This is partially explained by the limnology, shape and depth of Lake Ontario. 

Let's also explore 
```{r hp plots}
sample_and_taxon %>% 
ggplot()+
  aes(x = env_group,
      y = cells_per_ml)+
  geom_boxplot(aes(fill = env_group))+
  scale_fill_hp_d(option = "LunaLovegood", name = "Depth")+
  labs(x = "Environmental Group",
       y = "Microbial Abundance",
       title = "Impact of Environmental Group on Microbial Abundance")+
  stat_compare_means(method = "anova", label.y = 40)+     
  stat_compare_means(label = "p.signif", method = "t.test",ref.group = "0.5")
```

There appears to be significant differences between environmental groups.

This last data summary is a table looking at taxon Chloroflexi by environmental groups.

```{r figures}
sample_and_taxon %>% 
  group_by(env_group) %>% 
  summarise(avg_abundance = mean(Chloroflexi),
            std_dev = sd(Chloroflexi, na.rm = TRUE)) %>% 
  kable()

```








