---
title: "Lake Ontario Microbial Report"
author: "Riley Harding"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In this report we will investigate numerous dependent variables effecting microbial abundance in Lake Ontario.

# Set-up

#### Loading packages

First we load necessary packages:

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)
```

#### Reading in data

Then we will read in our data. We produced this dataframe by merging two dataframes together with an inner join base function in R.

```{r data, warning=FALSE, message=FALSE}
sample_and_taxon<-read_csv("~/OneDrive/Documents/Cornell Grad School/Graduate Classes/Workshops/2025_Carpentries_Coding/ontario-report/data/sample_and_taxon.csv")
```

# Analyzing the Data

Finally we will plot the relationship between temperature and overall microbial cell abundance.

```{r cells vs temp}
sample_and_taxon %>% 
  ggplot()+
  aes(x = temperature,
      y = cells_per_ml/1000000,
      color = env_group)+
  geom_point()+
  labs(x= "Temperature (°C)", y = "Cells (millions/mL)", title = "Are temperature and cell abundance linked?")
```

It appears the warmer waters promote higher microbial cell counts. Cell counts were especially high in shallow waters during September.

```{r data_summary, include=FALSE}
nSamples <- nrow(sample_and_taxon)

minTemp <- sample_and_taxon %>% 
  pull(temperature) %>% 
  min() %>% 
  round()

maxTemp <- sample_and_taxon %>% 
  pull(temperature) %>% 
  max() %>% 
  round()

```

The above plot shows `r nSamples` samples. For these samples the minimum temperature was `r minTemp`°C and the maximum temperature was `r maxTemp`°C.

| Statistic           | Value        |
|---------------------|--------------|
| Number of Samples   | `r nSamples` |
| Minimum Temperature | `r minTemp`  |
| Maximum Temperature | `r maxTemp`  |


```{r kable}
library(knitr)

sample_and_taxon %>% 
  filter(env_group == "Deep") %>% 
  select(cells_per_ml, temperature, Chloroflexi) %>% 
  kable()
```

Challenge: Make a bullet point list.
First line is your favorite coding language (italicize)
Second line is your favorite function (bold)
Third line is the next programming thing to learn (italics and bold)
Then a numbered list
Find Rmd formatting tutorial and insert as a hyperlink

* *R is my favorite coding language*
* **My favorite R function is facet within tidyverse package**
* ***The next thing I want to learn is how to collaborate with my PI via GitHub***


1. *R is my favorite coding language*
2. **My favorite R function is facet within tidyverse package**
3. ***The next thing I want to learn is how to collaborate with my PI via GitHub***

[Click here for a tutorial](https://rmarkdown.rstudio.com/authoring_basics.html)









